Three Algorithms for Merging Navigable Graphs

### 2024-10-08 paper/main.tex
add pseudocode for naive-merge alg
add text description of general, naive, and merge1 algorihtms

\usepackage{natbib}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}
\usepackage{amsmath}
\DeclareMathOperator*{\argminA}{arg\,min} % Jan Hlavacek
\DeclareMathOperator*{\argminB}{argmin}   % Jan Hlavacek
\usepackage{amsfonts}
\usepackage{amssymb}
\begin{document}

@@

We start from the basic search.

@@

\subsection{HNSW Merge Basic}

@@

% \State $\mathcal{H}_c \gets \emptyset$

@@

\subsection{Merge1}

@@

        \State $\mathcal{C}_{not\_done} \gets \{\mathcal{C}^b_1, \mathcal{C}^b_2, ..., \mathcal{C}^b_{ \text{next\_step\_k} } \} \cap \mathcal{V}_{not\_done}$

        \State $v^* \gets \underset{v \in \mathcal{C}_{not\_done}}{\mathrm{argmin}} \rho(v, v^*)$
    \EndWhile
\EndWhile

@@

>>>>

\usepackage{natbib}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}
\usepackage{amsmath}
\DeclareMathOperator*{\argminA}{arg\,min} % Jan Hlavacek
\DeclareMathOperator*{\argminB}{argmin}   % Jan Hlavacek
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{nameref}
\begin{document}

@@

We start from the basic search over the graph.

@@

We provide 3 merge algorithms: "Naive Merge", "Merge1", and "Merge2". We also refer to these algorithm as layer-merge algorithms. "Naive Merge" is a straight forward method, while "Merge1" and "Merge2" are more advanced. Together, these two algorithms constitute the main contribution of the work. The HNSW structure is a sequence of graphs, called layers. We start from description of a general algorithm for merging multi layers structure.

\subsection{HNSW General Merge}

HNSW-General-Merge algorithm \ref{alg:general_merge} is a straight forward. It uses one of the layer-merge algorithm to merge each HNSW layer. To able layer-merge algorithm to perform fast search using multi-layer structure of HNSW, HNSW-General-Merge algorithm passes to layer merge algorithm both HNSW structures $\mathcal{H}_a, \mathcal{H}_b$, and the layer number $\ell=i$ to which the merge should be performed. An order of merging layers doesn't mater. Thus merging of layers can be done in parallel.

@@

@@

\subsection{Naive Merge}
The Naive Merge algorithm \ref{alg:merge_naive} surprisingly is very simple. It takes one vertex $v^*$ by one from the graph $G^a$. Then algorithm prepares the candidate set $\mathcal{C}$ for a new neighborhood (lines 7,8). It uses \textsc{HNSW-Search} to get $m$-closest vertices to $v^*$ in the graph $G^b$, and combines with an old neighborhood of the vertex $v^*$. After that in the line 9 the candidate set $\mathcal{C}$ is passed into the neighborhood construction function, which selects the vertices for the new merged neighborhood for $v^*$.
Lines 10-13 contains the same logic for the forming new neigborhoods of graph $G^b$

\begin{algorithm}
\caption{\textsc{Merge-Naive}($\mathcal{H}_a, \mathcal{H}_b,  \ell, \text{search\_ef}$)}\label{alg:merge_naive}
\textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i)$, $\mathcal{H}_b = (G^b_i)$, and the level number $\ell$, search parameter $\text{search\_ef}$ \\
\textbf{Ensure:} Merged graph $G^c$ 
\begin{algorithmic}[1]
\State $V^c \gets V^a \cup V^b$
\State $E^c \gets \emptyset$
\State $G^a \gets G^a_{\ell}$
\State $G^b \gets G^b_{\ell}$
\State $m \gets \mathcal{H}_a.m_0$ if $\ell = 0$ else $\mathcal{H}_a.m$

\For{$v^* \in V^a$}
    \State $\mathcal{C}^b \gets \textsc{HNSW-Search}(\mathcal{H} = \mathcal{H}^b, q = v^*, v_0, k = m, L = \text{search\_ef}, \ell)$
    \State $\mathcal{C} \gets \{v : (v^*, v) \in E^a \} \cup \mathcal{C}^b$
    \State $E^c \gets E^c \cup \{(v^*, v) : v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$
\EndFor

\For{$v^* \in V^b$}
    \State $\mathcal{C}^a \gets \textsc{HNSW-Search}(\mathcal{H} = \mathcal{H}^a, q = v^*, v_0, k = m, L = \text{search\_ef}, \ell)$
    \State $\mathcal{C} \gets \{v : (v^*, v) \in E^b \} \cup \mathcal{C}^a$
    \State $E^c \gets E^c \cup \{(v^*, v) : v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$
\EndFor

\State \Return $G^c = (V^c, E^c)$
\end{algorithmic}
\end{algorithm}

\subsection{Merge1}
The most effort of \textsc{Merge-Naive} algorithm lies in the set of neighborhood candidates from the opposite graph utilising \textsc{HNSW-Search} procedure, that traverses layers graph from the top-level down to layer number $\ell$ every time. The number of computations can be reduced if we select the next vertex $v^*$ close to the previous one (line 15), instead of choosing it randomly. Thus, for new $v^*$ the neighborhood candidates will be  also close to the previous candidates set. To search this new neighborhood candidates we can use \textsc{LocalSearch} procedure, that traversing the same graph staring from the previous neighborhood candidates set $\mathcal{P}^b$ (line 11,14). In the line 14 in set $\mathcal{P}^b$ we keep only $M$-closest to $v^*$ candidates.\\
In the line 15 almost at each iteration we try select new $v^*$ vertex close to the previous  $v^*$ that was not already processed.  
To ensure that new $v^*$ vertex are not very far from the previous $v^*$ we bound the size of results of \textsc{LocalSearch} procedure controlling by "next\_step\_k" parameter. Once \textsc{LocalSearch} can't find enough close not processed vertex, we choice new $v^*$ from the set $\mathcal{V}_not\_done$ randomly.

@@

        \State $\mathcal{P}^b \gets \{\mathcal{C}^b_1, \mathcal{C}^b_2, ..., \mathcal{C}^b_M \} $

        \State $\mathcal{C}^a  \gets \textsc{LocalSearch}(G=G^a, q=v^*, C=\{v^*\} , k=\text{next\_step\_k}, L=\text{next\_step\_ef})$

        \State $\mathcal{C}^a \gets \mathcal{C}^a \cap \mathcal{V}_{not\_done}$

        \If{$\mathcal{C}^a = \emptyset$}
            \State \textbf{break}
        \EndIf

        \State $v^* \gets \mathcal{C}^a_1$
    \EndWhile
\EndWhile

@@

\While{$\mathcal{V}_{not\_done} \neq \emptyset$}
    \State $v^* \gets \text{randomly choice from } \mathcal{V}_{not\_done}$
